Quantity discrimination, decision-making, and the role of early-life conditions in a lizard

### 2025-05-20 ms.qmd
updates

---
title: "Quantity discrimination, decision-making, and the role of early-life conditions in a lizard"
docx: 
  reference-doc: "./bib/tmpl.docx"
bibliography: "./bib/refs.bib"
csl: "./bib/behavioral-ecology-and-sociobiology.csl"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

@@

## Abstract
Quantity discrimination affects a range of behaviours essential for fitness, including social interactions, navigation, and foraging. Animals use different cognitive systems to distinguish between quantities. The Approximate Number System (ANS) estimates, while the Object File System (OFS) tracks discrete items. In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested. Early life conditions can also influence brain development and cognition, potentially affecting quantity discrimination and decision-making. Here, we tested the ability of the common garden skink (_Lampropholis guichenoti_) to discriminate between quantities and the effects of elevated prenatal GCs and incubation temperature on quantity discrimination and decision-making. We assessed quantity discrimination using a spontaneous choice test with food as a stimulus. We subjected lizards to five two-choice tests differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4). Contrary to our predictions, we found no evidence of quantity discrimination in _L. guichenoti_. Furthermore, prenatal conditions did not influence their performance or decision-making. Despite these results, further investigation of quantity discrimination in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.

>>>>

---
title: "Quantity discrimination, decision-making, and the role of early-life conditions in a lizard"
docx: 
  reference-doc: "./bib/tmpl.docx"
bibliography: "./bib/refs.bib"
csl: "./bib/animal-cognition.csl"
execute:
  echo: false
  error: false
  cache: false
  warning: false
link-citations: true
crossref:  
  fig-title: Fig    # (default is "Figure")
  title-delim: —     # (default is ":")
  fig-prefix: Fig.   # (default is "Figure")
  tbl-prefix: Table  # (default is "Table")
editor_options: 
  chunk_output_type: console
---

@@

## Abstract
Quantity discrimination affects a range of behaviours essential for fitness, including social interactions, navigation, and foraging. Animals use different cognitive systems to distinguish between quantities. The Approximate Number System (ANS) estimates, while the Object File System (OFS) tracks discrete items. In reptiles, quantity discrimination remained largely unexplored until recently, and the use of one system over the other can depend on the species or the context in which they are tested. Early life conditions can also influence brain development and cognition, potentially affecting quantity discrimination and decision-making. Here, we tested the ability of the common garden skink (_Lampropholis guichenoti_) to discriminate between quantities and the effects of elevated prenatal GCs and incubation temperature on quantity discrimination and decision-making. We assessed quantity discrimination using a spontaneous choice test with food as a stimulus. We subjected lizards to five two-choice tests differing in the number of items (1 VS 4, 1 VS 3, 2 VS 4, 2 VS 3, 3 VS 4) while controlling for total length and area. Contrary to our predictions, we found no evidence of quantity discrimination in _L. guichenoti_. Furthermore, prenatal conditions did not influence their performance or decision-making. Despite these results, further investigation of quantity discrimination in live-prey-feeding reptiles and the impact of early conditions on other taxa remains an important area of research.
